Preliminary Draft

### 2021-07-20 article/draft.tex
updates

Furthermore, mainstream platforms are increasingly engaging in editorial tasks by implementing targeted policies to insure that each platform's rules are not violated. Community guidelines of Facebook, Twitter and Youtube can be summarized in a handful of categories, regarding safety, privacy and authenticity; which include sub-categories such as violence, terrorism, child sexual exploitation, abuse, harassment, hateful conduct, suicide or self-harm, illegal or regulated goods and services, platform manipulation and spam (see table \ref{summary} in Appendix \ref{links}). While specific to each platform, the previously cited categories correspond in most cases to well defined concepts that fall into legal frameworks in many countries, unlike misinformation. The intricacies of constructing a legal framework for misinformation arises from the difficulty of identifying and qualifying a piece of online content as false or misleading, among an overwhelming quantity of daily produced content, without infringing existing laws.\footnote{See \href{https://www.poynter.org/ifcn/anti-misinformation-actions/}{A guide to anti-misinformation actions around the world} on the website of Poynter Institute.} In particular, a number of recent studies  point towards the idea that ``Fake News'' or disinformation is a small subset of the total supply of information on online social networking platforms (e.g. Grinberg et al. (2019)~\cite{grinberg} and Broniatowski et al. (2020)~\cite{broniatowski}). Yet, this seemingly small subset is generating great concern in traditional media and in society in a broader sense. \footnote{For example see the February 2020 \href{https://www.who.int/director-general/speeches/detail/munich-security-conference}{speech} of the Director General of the WHO at the Munich Security Conference, where he says ``But we’re not just fighting an epidemic; we’re fighting an infodemic.'' Furthermore, the European commission recognizes the spread of online disinformation as a problem and has put together in 2018 a \href{https://digital-strategy.ec.europa.eu/en/policies/code-practice-disinformation}{Code of practice on Disinformation}, which is a set of self-regulatory standards to fight disinformation. {\color{pink} Add more examples to show that this problem is highlighted by different institutions with different/neutral political color.. }}

@@

In this article, we focus on mainstream platforms' policies and actions regarding content with low credibility or false information, commonly referred to as {\it Fake News} (for an overview of the concept see Lazer et al. (2018)~\cite{lazer}). The {\it Fake News} phenomenon is still ill-defined by the academic community, as it encompasses several combined features such as spreading inaccurate, false or misleading information, with or without the intention of influencing or manipulating a target pool of audience.  The growth of social networking platforms over the last decade in terms of number of users worldwide and volume of content, has modified the information ecosystem in terms of production of information and its mediation. Many users can now produce and share content which includes news related information, without having to abide by strict editorial processes that ensure accuracy of information and reliability of sources. In particular, false or inaccurate content produced and shared on social networking platforms concerning the political life or public health may have a potentially harmful impact on the society, in the rare event that it goes viral. This gave rise to a set of heterogenous fact-checking policies across mainstream platforms. For example, Facebook has a substantial partnership program with Fact-checking partners certified by the non-partisan International Fact-Checking Network. Facebook use a number of signals and machine learning models to predict misinformation and surface it to fact-checkers.\footnote{See the \href{https://www.facebook.com/journalismproject/programs/third-party-fact-checking/how-it-works}{section} Frequently asked questions: `How does Facebook use technology to detect potential misinformation?''} Twitter seems to have a different approach where they focus on providing context rather than fact-checking\footnote{To the best of our knowledge, Twitter does not have a page which summarizes its fact-checking strategy. The Twitter Safety Team tweeted on June 3, 2020 the following:  ``We heard: 1. Twitter shouldn’t determine the truthfulness of Tweets 2. Twitter should provide context to help people make up their own minds in cases where the substance of a Tweet is disputed. Hence, our focus is on providing context, not fact-checking.'' Tweet ID \href{https://twitter.com/TwitterSafety/status/1267986503721988096}{1267986503721988096}.} and the platform is testing a new system based on the wisdom of the crowds to tackle misinformation (see \href{https://blog.twitter.com/en\_us/topics/product/2021/introducing-birdwatch-a-community-based-approach-to-misinformation}{Twitter Birdwatch}). As for Youtube, this platform utilizes the \href{https://schema.org/ClaimReview}{schema.org ClaimReview} markup, where fact-checking articles created by eligible publishers can appear on information panels (see table \ref{summary} in Appendix \ref{links}).

@@

During the COVID19 global health pandemic platforms have upgraded their guidelines to include a set of rules to tackle the propagation of potentially harmful content (see table \ref{summary} in Appendix \ref{links}). Those policies are enforced via existing actions used by the platforms to tackle other rules' violations, such as: labelling content to provide more context or indicate falsehood, publishing a list of terms or topics that will be flagged\footnote{{\color{pink} put here the list of Facebook and Twitter terms about covid.}}, suspending accounts, implementing strike systems, reducing the visibility of content, etc. As each platform is a private company, those {\it new} policies are not coordinated and are implemented in different ways across platforms. Such targeted policies show the willingness of mainstream platforms to enhance the quality of the online conversation, but also sheds light on the lack of specific policies to tackle misinformation in general. In particular, policies regarding misinformation are not part of the set of platform rules or community guidelines\footnote{As of July 2021.}. The 2019 report of the Facebook Data Transparency Advisory Group\footnote{To read the full report, see: \href{https://law.yale.edu/yls-today/news/facebook-data-transparency-advisory-group-releases-final-report?fbclid=IwAR2xMZr5GdD1GaNpjsXR3\_yeeIR4H9iFASfrni5HKcJVAO5oWA52bvwcZxU}{https://law.yale.edu/yls-today/news/facebook-data-transparency-advisory-group-releases-final-report?fbclid=IwAR2xMZr5GdD1GaNpjsXR3\_yeeIR4H9iFASfrni5HKcJVAO5oWA52bvwcZxU}.} (DTAG) states that ``{\it DTAG was not tasked with evaluating any of the following: (...) Facebook’s policies with respect to “fake news” or misinformation, as neither of these categories were counted as violations within the first two versions of the Community Standards Enforcement Report.}''. Facebook's strategy to tackle ``False News'' is three fold : Remove, Reduce, Inform. It is explained via a blog post on the facebook Newsroom.\footnote{See: \href{https://about.fb.com/news/2018/05/hard-questions-false-news/}{https://about.fb.com/news/2018/05/hard-questions-false-news/}.}. Similarly, Twitter communicates about actions related to misinformation via their Twitter Safety Blog/account\footnote{See \href{https://blog.twitter.com/en\_us/authors.TwitterSafety}{https://blog.twitter.com/en\_us/authors.TwitterSafety}.}. In a post on Youtube Official Blog\footnote{See \href{https://blog.youtube/inside-youtube/the-four-rs-of-responsibility-raise-and-reduce/}{https://blog.youtube/inside-youtube/the-four-rs-of-responsibility-raise-and-reduce/}.}, the platform explained its ``Four Rs of Responsibility" and how it raises authoritative content, reduces borderline content and harmful misinformation.

@@

Nevertheless, misinformation specific interventions by mainstream platforms are hard to monitor, study or verify by third parties (e.g. academic community, data journalists, NGOs), as in many cases misleading or false content does not qualify as a violation of a given platform's rules and it does not explicitly appear as a separate category in available transparency reports. This makes the study of online misinformation, the assessment of the impact of platforms' actions to tackle misinformation and their relevancy a burdensome task for the academic community. Hence, in the present article we explain how to verify with data mining mainstream platforms' current actions regarding content with low credibility or false information. We do so by providing a series of examples for different interventions and platforms. We only focus on three platforms: Facebook, Twitter and Youtube. Both Facebook and Youtube are in the top three most popular social media platforms in terms of number of users.\footnote{See for example the ranking of the most popular social networks as of April 2021 on Statista: \href{https://www.statista.com/statistics/272014/global-social-networks-ranked-by-number-of-users/}{https://www.statista.com/statistics/272014/global-social-networks-ranked-by-number-of-users/}.} We further choose Twitter because it is a social networking platform with the most news-focused users, according to the Pew Research center (2019)~\cite{pew1}. More specifically, we survey a number of common policies used in order to tackle misinformation, across the three above cited mainstream platforms, Facebook, Twitter and Youtube: temporary or permanent suspension of users, reducing the visibility of some content, introducing flags and notices. We do not provide  an exhaustive list of methods on how to investigate the platforms’ policies.\footnote{For a detailed list of interventions check the following \href{https://airtable.com/shrO0ooI9WSEfIUhb/tblAWQwFOiihKdQjm/viwZLAOzLK1NQ0c2n?blocks=hide}{airtable} compiled by Slatz and Leibowicz (2021)~\cite{niemanlab}. } We rather provide a methodology to investigate key policies, that can be useful to researchers or journalists interested in implementing external monitoring. We summarize in table \ref{tab1} the tools used to collect the data from Facebook, Twitter and Youtube, that we use in multiple examples throughout the present article. We compile in Appendix \ref{links} in table \ref{summary}, URL links that redirect to the policies, regulations and transparency centers of Facebook, Twitter and Youtube. Finally, we discuss how an increased effort of transparency regarding specific content can help the community of researchers study and assess the impact of platforms' policies regarding misinformation.  %the announced policy for a specific user was implemented or not.

@@

%paragraph strike system + say following example is a manual decision. 
Twitter has implemented a strike system as part of their Civi Integrity Policy\footnote{See \href{https://help.twitter.com/en/rules-and-policies/election-integrity-policy}{help.twitter.com/en/rules-and-policies/election-integrity-policy}} and their COVID19 misleading information policy.\footnote{See \href{https://help.twitter.com/en/rules-and-policies/medical-misinformation-policy}{help.twitter.com/en/rules-and-policies/medical-misinformation-policy} and \href{https://blog.twitter.com/en\_us/topics/company/2021/updates-to-our-work-on-covid-19-vaccine-misinformation}{blog.twitter.com/en\_us/topics/company/2021/updates-to-our-work-on-covid-19-vaccine-misinformation}} Violations of both policies can entail strikes, where two strikes lead to a 12-hour account lock and five or more strikes lead to permanent suspension from the platform. The 12-hour account lock is hard to observe in the data, especially for users who do not have an over the clock tweeting activity. In this section, we provide one example of a temporary suspension\footnote{See the official documentation on the Twitter's Help Center regarding account suspension: \href{https://help.twitter.com/en/managing-your-account/suspended-twitter-accounts}{https://help.twitter.com/en/managing-your-account/suspended-twitter-accounts }.} of a Twitter account, that seems to be the result of a manual decision concerning a Tweet which violated the rules.

@@

	\item Finally, we found irregularities in the number of Infowars articles collected from Buzzsumo. While Infowars usually publishes $20$ to $30$ articles per day, only $53$ articles were collected in the $31$-day period between June $11$ to July $11$, 2020. A temporary crawling problem coming from Buzzsumo may have caused this lack of data. Because no database is perfect, we would like to highlight the importance of cross-checking information between different sources when possible. 
\end{itemize}
}

@@

\multirow{3}{*}{Rules}                     & \includegraphics[scale=0.05]{./img/fb_logo.png} & \href{https://www.facebook.com/communitystandards/recentupdates/}{facebook.com/communitystandards/recentupdates/}                         \\ \cline{2-3} 

@@

\multirow{3}{*}{\begin{tabular}[c]{@{}cl@{}}  Transparency \\ center \end{tabular}}       & \includegraphics[scale=0.05]{./img/fb_logo.png} & https://transparency.fb.com/data/                                                  \\ \cline{2-3} 

@@

 \end{tabular}
\caption{Summary of ressources, last accessed on July 5, 2021. }
\label{summary}
\end{table*}

>>>>

Furthermore, mainstream platforms are increasingly engaging in editorial tasks by implementing targeted policies to insure that each platform's rules are not violated. Community guidelines of Facebook, Twitter and Youtube can be summarized in a handful of categories, regarding safety, privacy and authenticity; which include sub-categories such as violence, terrorism, child sexual exploitation, abuse, harassment, hateful conduct, suicide or self-harm, illegal or regulated goods and services, platform manipulation and spam (see table \ref{summary} in Appendix \ref{links}). While specific to each platform, the previously cited categories correspond in most cases to well defined concepts that fall into legal frameworks in many countries, unlike misinformation. The intricacies of constructing a legal framework for misinformation arises from the difficulty of identifying and qualifying a piece of online content as false or misleading, among an overwhelming quantity of daily produced content, without infringing existing laws.\footnote{See \href{https://www.poynter.org/ifcn/anti-misinformation-actions/}{A guide to anti-misinformation actions around the world} on the website of Poynter Institute.} In particular, a number of recent studies  point towards the idea that ``Fake News'' or disinformation is a small subset of the total supply of information on online social networking platforms (e.g. Grinberg et al. (2019)~\cite{grinberg} and Broniatowski et al. (2020)~\cite{broniatowski}). Yet, this seemingly small subset is generating great concern in traditional media and in society in a broader sense. \footnote{For example see the February 2020 \href{https://www.who.int/director-general/speeches/detail/munich-security-conference}{speech} of the Director General of the WHO at the Munich Security Conference, where he says ``But we’re not just fighting an epidemic; we’re fighting an infodemic.'' Furthermore, the European commission recognizes the spread of online disinformation as a problem and has put together in 2018 a \href{https://digital-strategy.ec.europa.eu/en/policies/code-practice-disinformation}{Code of practice on Disinformation}, which is a set of self-regulatory standards to fight disinformation.}

@@

Hence, in the present article, we focus on mainstream platforms' policies and interventions regarding content with low credibility or false information, commonly referred to as {\it Fake News} (see Lazer et al. (2018)~\cite{lazer}). The {\it Fake News} phenomenon is still ill-defined by the academic community, as it encompasses several combined features such as spreading inaccurate, false or misleading information, with or without the intention of influencing or manipulating a target pool of audience.  The growth of social networking platforms over the last decade in terms of number of users worldwide and volume of content, has modified the information ecosystem in terms of production of information and its mediation. Many users can now produce and share content which includes news related information, without having to abide by strict editorial processes that ensure accuracy of information and reliability of sources. In particular, false or inaccurate content produced and shared on social networking platforms concerning the political life or public health may have a potentially harmful impact on the society, in the rare event that it goes viral. This gave rise to a set of heterogenous fact-checking policies across mainstream platforms. For example, Facebook has a substantial partnership program with Fact-checking partners certified by the non-partisan International Fact-Checking Network. Facebook uses a number of signals and machine learning models to predict misinformation and surface it to fact-checkers.\footnote{See the \href{https://www.facebook.com/journalismproject/programs/third-party-fact-checking/how-it-works}{section} Frequently asked questions: `How does Facebook use technology to detect potential misinformation?''} Twitter seems to have a different approach where they focus on providing context rather than fact-checking\footnote{To the best of our knowledge, Twitter does not have a page which summarizes its fact-checking strategy. The Twitter Safety Team tweeted on June 3, 2020 the following:  ``We heard: 1. Twitter shouldn’t determine the truthfulness of Tweets 2. Twitter should provide context to help people make up their own minds in cases where the substance of a Tweet is disputed. Hence, our focus is on providing context, not fact-checking.'' Tweet ID \href{https://twitter.com/TwitterSafety/status/1267986503721988096}{1267986503721988096}.} and the platform is testing a new system based on the wisdom of the crowds to tackle misinformation (see \href{https://blog.twitter.com/en\_us/topics/product/2021/introducing-birdwatch-a-community-based-approach-to-misinformation}{Twitter Birdwatch}). As for Youtube, this platform utilizes the \href{https://schema.org/ClaimReview}{schema.org ClaimReview} markup, where fact-checking articles created by eligible publishers can appear on information panels (see table \ref{summary} in Appendix \ref{links}).

@@

During the COVID19 global health pandemic platforms have upgraded their guidelines to include a set of rules to tackle the propagation of potentially harmful content (see table \ref{summary} in Appendix \ref{links}). Those policies are enforced via existing actions used by the platforms to tackle other rules' violations, such as: labelling content to provide more context or indicate falsehood, publishing a list of terms or topics that will be flagged\footnote{{\color{pink} put here the list of Facebook and Twitter terms about covid.}}, suspending accounts, implementing strike systems, reducing the visibility of content, etc. As each platform is a private company, those {\it new} policies are not coordinated and are implemented in different ways across platforms. Such targeted policies show the willingness of mainstream platforms to enhance the quality of the online conversation, but also sheds light on the lack of specific policies to tackle misinformation in general. In particular, policies regarding misinformation are not part of the set of platform rules or community guidelines\footnote{As of July 2021.}. The 2019 report of the Facebook Data Transparency Advisory Group (DTAG) states that ``{\it DTAG was not tasked with evaluating any of the following: (...) Facebook’s policies with respect to “fake news” or misinformation, as neither of these categories were counted as violations within the first two versions of the Community Standards Enforcement Report.}''. Facebook's strategy to tackle ``False News'' is three fold : Remove, Reduce, Inform. It is explained via a blog post on the facebook Newsroom. Similarly, Twitter communicates about actions related to misinformation via their \href{https://blog.twitter.com/en\_us/authors.TwitterSafety}{Twitter Safety} Blog/account. In a post on Youtube Official Blog, the platform explained its ``Four Rs of Responsibility" and how it raises authoritative content, reduces borderline content and harmful misinformation (see table \ref{summary} in Appendix \ref{links}).

@@

Nevertheless, misinformation specific interventions by mainstream platforms are hard to monitor, study or verify by third parties (e.g. academic community, data journalists, NGOs), as in many cases misleading or false content does not qualify as a violation of a given platform's rules and it does not explicitly appear as a separate category in available transparency reports. This makes the study of online misinformation, the assessment of the impact of platforms' actions to tackle misinformation and their relevancy a burdensome task for the academic community. Hence, in the present article we explain how to verify with data mining mainstream platforms' current actions regarding content with low credibility or false information. We do so by providing a series of examples for different interventions and platforms. We only focus on three platforms: Facebook, Twitter and Youtube. Both Facebook and Youtube are in the top three most popular social media platforms in terms of number of users.\footnote{See for example the ranking of the most popular social networks as of April 2021 on Statista: \href{https://www.statista.com/statistics/272014/global-social-networks-ranked-by-number-of-users/}{https://www.statista.com/statistics/272014/global-social-networks-ranked-by-number-of-users/}.} We further choose Twitter because it is a social networking platform with the most news-focused users, according to the Pew Research center (2019)~\cite{pew1}. More specifically, we survey a number of common policies used in order to tackle misinformation, across the three above cited mainstream platforms, Facebook, Twitter and Youtube: temporary or permanent suspension of users, reducing the visibility of some content, introducing flags and notices. We do not provide  an exhaustive list of methods on how to investigate the platforms’ policies.\footnote{For a detailed list of interventions check the following \href{https://airtable.com/shrO0ooI9WSEfIUhb/tblAWQwFOiihKdQjm/viwZLAOzLK1NQ0c2n?blocks=hide}{airtable} compiled by Slatz and Leibowicz (2021)~\cite{niemanlab}. } We rather provide a methodology to investigate key policies, that can be useful to researchers or journalists interested in implementing external monitoring. We summarize in table \ref{tab1} the tools used to collect the data from Facebook, Twitter and Youtube, that we use in multiple examples throughout the present article. We compile in Appendix \ref{links} in table \ref{summary}, a list of links that redirect to the policies, regulations and transparency centers of Facebook, Twitter and Youtube that we discuss throughout the article. Finally, we discuss how an increased effort of transparency regarding specific content can help the community of researchers study and assess the impact of platforms' policies regarding misinformation.  %the announced policy for a specific user was implemented or not.

@@

%paragraph strike system + say following example is a manual decision. 
Twitter has implemented a strike system as part of their Civic Integrity Policy and their COVID19 misleading information policy. Violations of both policies can entail strikes, where two strikes lead to a 12-hour account lock and five or more strikes lead to permanent suspension from the platform. The 12-hour account lock is hard to observe in the data, especially for users who do not have an over the clock tweeting activity. In this section, we provide one example of a temporary suspension\footnote{See the official documentation on the Twitter's Help Center regarding account suspension: \href{https://help.twitter.com/en/managing-your-account/suspended-twitter-accounts}{https://help.twitter.com/en/managing-your-account/suspended-twitter-accounts }.} of a Twitter account, that seems to be the result of a manual decision concerning a Tweet which violated the rules.

@@

	\item Finally, we found irregularities in the number of Infowars articles collected from Buzzsumo. While Infowars usually publishes $20$ to $30$ articles per day, only $53$ articles were collected in the $31$-day period between June $11$ to July $11$, 2020. A temporary crawling problem coming from Buzzsumo may have caused this lack of data. Because no database is perfect, we would like to highlight the importance of cross-checking information between different sources when possible.

	\item We hope this paper can illustrate how the platforms policies can be verified. All the code needed to collect the data, and create the figures can be found on a public GitHub repo: https://github.com/medialab/truth-and-trust-online-2021/, so that other researchers and journalists can easily reuse our methodology. 
\end{itemize}
}

@@

\multirow{3}{*}{Rules}                     & \includegraphics[scale=0.05]{./img/fb_logo.png} & \begin{tabular}[c]{@{}cl@{}}   \href{https://www.facebook.com/communitystandards/recentupdates/}{facebook.com/communitystandards/recentupdates/}    \end{tabular}                     \\ \cline{2-3} 

@@

\multirow{3}{*}{\begin{tabular}[c]{@{}cl@{}}  Transparency \\ center \end{tabular}}       & \includegraphics[scale=0.05]{./img/fb_logo.png} & \begin{tabular}[c]{@{}cl@{}}  \href{https://transparency.fb.com/data/}{transparency.fb.com/data/}    \\ \href{https://law.yale.edu/yls-today/news/facebook-data-transparency-advisory-group-releases-final-report}{https://law.yale.edu/yls-today/news/facebook-data-transparency-advisory-group-releases-final-report} \end{tabular}                                               \\ \cline{2-3} 

@@

%

\multirow{3}{*}{\begin{tabular}[c]{@{}cl@{}} Strike System \end{tabular}} & \includegraphics[scale=0.05]{./img/fb_logo.png} & \begin{tabular}[c]{@{}cl@{}} See in the following link the section {\it What is the number of strikes a person or } \\ { \it Page has to get to before you ban them?} \\  \href{https://about.fb.com/news/2018/08/enforcing-our-community-standards/}{about.fb.com/news/2018/08/enforcing-our-community-standards/}   \end{tabular}                            \\ \cline{2-3} 
                                           & \includegraphics[scale=0.007]{./img/tw_logo.png}   &  \begin{tabular}[c]{@{}cl@{}}  See in the following links the section:  {\it Account locks and permanent suspension }   \\ \href{https://help.twitter.com/en/rules-and-policies/election-integrity-policy}{help.twitter.com/en/rules-and-policies/election-integrity-policy}  \\ \href{https://help.twitter.com/en/rules-and-policies/medical-misinformation-policy}{help.twitter.com/en/rules-and-policies/medical-misinformation-policy}  \end{tabular}    \\ \cline{2-3} 
                                           & \includegraphics[scale=0.03]{./img/yt_logo.png}  & \begin{tabular}[c]{@{}cl@{}}    \end{tabular}                            \\ \hline 
 \end{tabular}
\caption{Summary of ressources, last accessed on July 5, 2021. }
\label{summary}
\end{table*}
